---
title: "plots"
author: "NM"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(here)
library(tmap)
library(ggplot2)
library(sf)
library(terra)
library(raster)
library(tidycensus)
library(data.table)
```

```{r}
# Visualize temperature data distribution
santaclara_temp_sj <- santaclara_temp %>%
  filter(station == 'San Jose')

santaclara_temp_rh <- santaclara_temp %>%
  filter(station == 'Reid-Hill')

santaclara_temp_m <- santaclara_temp %>%
  filter(station == 'Moffett')

sj_temp_dist <- ggplot(data = santaclara_temp_sj, aes(x = max_temp))+
  geom_histogram(binwidth = 2)

rh_temp_dist <- ggplot(data = santaclara_temp_rh, aes(x = max_temp))+
  geom_histogram(binwidth = 2)

m_temp_dist <- ggplot(data = santaclara_temp_m, aes(x = max_temp))+
  geom_histogram(binwidth = 2)

sj_temp_dist
rh_temp_dist
m_temp_dist
```

```{r}
# Difference in proportions of days over 90?
hot_days_prop_plot <- ggplot(data = santaclara_temp, aes(station, fill = factor(over_90)))+
  geom_bar() +
 scale_fill_brewer("Temp >= 90", palette = "Dark2")

hot_days_prop_plot
```

Use chi-square goodness of fit test to determine if frequency of number of days over 90 degrees Celsius are different to our expected null hypothesis. We want to test a hypothesis about 

H0: The proportion of days of 90 degrees Celsius or higher, measured at San Jose, Moffett and Reid-Hill from 2009 - 2023, are the same.
HA: The proportion of days of 90 degrees Celsius or higher, measured at San Jose, Moffett and Reid-Hill from 2009 - 2023, are different.

```{r}
# Create a data frame from our larger table
chi_table_clean <- santaclara_temp %>%
  subset(., select = c(station, over_90))

#chi_table <- data.frame(station=c('San Jose', 'Reid_Hill', 'Moffett'),
 #                            over_90=c(219, 430, 83),
  #                           under_90=c(5259, 4937, 4590))

chi_table_test <- table(chi_table_clean$station, chi_table_clean$over_90)

print(chi_table_test)
```

```{r}
# Apply chisq.test()
chisq.test(chi_table_test)
```

Take RH and Moffett and run a lm - to determine if means of 

```{r}
income_station_regression <- lm(formula = max_temp ~ mean_income, data = santaclara_temp)

summary(income_station_regression)
```

chi2 and then regression coefficient - ANOVA if time