---
title: "plots"
author: "NM"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(here)
library(tmap)
library(ggplot2)
library(sf)
library(terra)
library(raster)
library(tidycensus)
library(data.table)
library(patchwork)
```

```{r}
# Visualize temperature data distribution
santaclara_temp_sj <- santaclara_temp %>%
  filter(station == 'San Jose')

santaclara_temp_rh <- santaclara_temp %>%
  filter(station == 'Reid-Hill')

santaclara_temp_m <- santaclara_temp %>%
  filter(station == 'Moffett')

sj_temp_dist <- ggplot(data = santaclara_temp_sj, aes(x = max_temp))+
  geom_histogram(binwidth = 2, fill = 'cornflowerblue')+
  theme_bw() +
  ggtitle('Temps from San Jose Mineta Airport (2009 - 2023)')

rh_temp_dist <- ggplot(data = santaclara_temp_rh, aes(x = max_temp))+
  geom_histogram(binwidth = 2, fill = 'cornflowerblue') +
  theme_bw() +
  ggtitle('Temps from Reid-Hillview Airport (2009 - 2023)')

m_temp_dist <- ggplot(data = santaclara_temp_m, aes(x = max_temp))+
  geom_histogram(binwidth = 2, fill = 'cornflowerblue') +
  theme_bw() +
  ggtitle('Temps from Moffett Federal Airfield (2009 - 2023)')

sj_temp_dist /
rh_temp_dist /
m_temp_dist
```

```{r}
# Difference in proportions of days over 90?
hot_days_prop_plot <- ggplot(data = santaclara_temp, aes(station, fill = factor(over_90)))+
  geom_bar() +
 scale_fill_brewer("Temp >= 90", palette = "Paired") +
  ggtitle("Proportion of Days >= 90 Celsius")

hot_days_prop_plot
```

Use chi-square goodness of fit test to determine if frequency of number of days over 90 degrees Celsius are different to our expected null hypothesis. We want to test a hypothesis about 

H0: The proportion of days of 90 degrees Celsius or higher, measured at San Jose, Moffett and Reid-Hill from 2009 - 2023, are the same.
HA: The proportion of days of 90 degrees Celsius or higher, measured at San Jose, Moffett and Reid-Hill from 2009 - 2023, are different.

```{r}
# Create a data frame from our larger table
chi_table_clean <- santaclara_temp %>%
  subset(., select = c(station, over_90))

#chi_table <- data.frame(station=c('San Jose', 'Reid_Hill', 'Moffett'),
 #                            over_90=c(219, 430, 83),
  #                           under_90=c(5259, 4937, 4590))

chi_table_test <- table(chi_table_clean$station, chi_table_clean$over_90)

print(chi_table_test)
```

```{r}
# Apply chisq.test()
print(chisq.test(chi_table_test))
```

Take RH and Moffett and run a lm - to determine if means of 

```{r}
income_station_regression <- lm(formula = max_temp ~ mean_income + station, data = santaclara_temp)

summary(income_station_regression)
```

Does mean income have a relationship with station?

```{r}
ggplot(santaclara_temp, aes(station, mean_income))+
  geom_boxplot() +
  ggtitle("Median Income by Station")
```

H0: There is no difference between the mean incomes of reid-hill and moffett.
HA: There is a difference between the mean incomes of reid-hill and moffett.

```{r}
santaclara_income_station <- santaclara_temp %>% filter(station %in% c('Reid-Hill', 'Moffett'))

santaclara_income_station <- santaclara_income_station %>%
  group_by(station) %>%
  summarize(income_mean = mean(mean_income))

point_est_income_station <- santaclara_income_station$income_mean[2] - santaclara_income_station$income_mean[1]
```

```{r}
null_dist <- replicate(1000, {
  santaclara_income_station <- santaclara_income_station %>%
  mutate(station = sample(station, n())) %>%
  group_by(station)  %>%
  summarize(income_mean = mean(mean_income))

point_est_income_station <- santaclara_income_station$income_mean[2] - santaclara_income_station$income_mean[1]
})

ggplot(tibble(null_dist), aes(null_dist))+
  geom_histogram(bins = 1000,
                 color = "cornflowerblue",
                 fill = NA) +
  geom_vline(xintercept = point_est_income_station,
             color = 'firebrick')+
  ggtitle("Sample Statistic under Null Hypothesis")
```

```{r}
# Calculate p-value
sum(abs(null_dist) > abs(point_est_income_station))/
  length(null_dist)
```

Is there a relationship between median income and max_temp?

H0: The 
HA:

```{r}
santaclara_income_temp_lm <- lm(max_temp ~ mean_income,
                        santaclara_income_station)
summary(santaclara_income_temp_lm)
```

```{r}
beta1_est <- summary(santaclara_income_temp_lm)$coefficients[2,1]
beta1_se <- summary(santaclara_income_temp_lm)$coefficients[2,2]

tibble(beta1 = seq(-(beta1_est + beta1_se),
                   beta1_est + beta1_se,
                   length.out = 200),
       density = dnorm(beta1, mean = 0, sd = beta1_se)) %>%
  ggplot(aes(beta1, density)) +
  geom_line(color = "cornflowerblue")+
  geom_vline(xintercept =  beta1_est, color = "firebrick")
```

```{r}
# Calculate p-value
pval <- 2*pnorm(-abs(beta1_est), mean = 0, sd = beta1_se)
pval
```

